---
title: "ddPCR"
author: "J.Schuy"
date: "30 4 2021"
output: pdf_document
---

#### ddPCR Script 
The following script takes raw calculated copy number / µL from ddPCR analysed in Quanta Soft software as input.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setenv(LANG="en")


```


Built with  `r getRversion()`
```{r setupUser, include=TRUE, echo=FALSE}
Sys.setenv(LANG="en")
rm(list=ls())
options(scipen=999)

library(ggplot2)
library(dplyr)
library(readxl)
library(writexl)
library(car)

#Paramters
ProjectName <- "ddPCR8"
inputname <- "20210427-ddpcr8-setup.xlsx"
tabname <- "single"
colorsForLegend <- c("grey", "black", "dodgerblue", "black", "red", "green", "palegreen1", "palegreen4", "darkgoldenrod1", "darkorchid1")
thresholdForPval <- 0.05
SkipExportSection <- TRUE
ChosenPatient <- "ALI9"
target <- "jct1"
reference <- "rpp30"
```

 

```{r Load input and create work environment}
#Working directory
# setwd(paste(rstudioapi::getSourceEditorContext()$path, "/../..", sep = ""))

#create folder for output graphs
if (!dir.exists(paste(getwd(), "/Graphs_", ProjectName, "/", sep="")))  {
  dir.create(paste(getwd(), "/Graphs_", ProjectName, "/", sep=""))      }
#create folder for extracted data like lists and dataframes
if (!dir.exists(paste(getwd(), "/ExtractedData_", ProjectName, "/", sep=""))) {
  dir.create(paste(getwd(), "/ExtractedData_", ProjectName, "/", sep=""))     }

#load all samples
if (file.exists(paste(getwd(), "/ExtractedData_", ProjectName, "/saveEnvironment.RData", sep=""))) {
  load(paste(getwd(), "/ExtractedData_", ProjectName, "/saveEnvironment.RData", sep=""))} else {


#import data, genes for current Project (here: allChr)
data_raw <- read_excel(paste0(getwd(), "/../",inputname), tabname)

#clean input
data_clean <- data_raw[1:24,1:18]

#remove unnecessary columns
data_reduced <- data_clean[,c(1:4,6:8, 14)]



#save cleaned df
write_xlsx(data_reduced, paste0(getwd(),"/ExtractedData_", ProjectName, "/raw_cleaned.xlsx"))

#clean env
rm(data_raw)

save.image(file = paste(getwd(), "/ExtractedData_", ProjectName, "/saveEnvironment.RData", sep=""))

}

#show
head(data_reduced)

```



## Normalise
Reference genes is avaraged per sample, then the conc of target is divided by that ref value yielding one normalized fraction of amplified structure per input. 
Here: biolg. triplicates -> three values

A second step includes averaging these three values per sample, returns mean and sd

```{r Normal over reference gene}

## normalize ----
PlotName <- "ddPCR8"

#get backup
dd8 <- data_reduced

#derive reference value for each sample
dd8_mean <- dd8 %>% 
  group_by(Sample, .add = TRUE) %>%
  group_by(Target, .add = TRUE) %>%
  summarise(mean = mean(`Conc(copies/µL)`), 
            SD = sd(`Conc(copies/µL)`))

#normalize each target with respective ref value
dd8_target <- dd8[dd8$Target == target,]

##loop for normalization
dd8_norm <- vector()
for (i in levels(factor(dd8_target$Sample))){

  subdat <- dd8_target[dd8_target$Sample == i,]
  
  #add ref value
  subdat$refmean <- dd8_mean$mean[dd8_mean$Target == reference &
                                  dd8_mean$Sample == i] 

  #normalize for "i"
  subdat$concNorm <- subdat$`Conc(copies/µL)` / subdat$refmean
  
  #mean and sd for plot
  subdat$mean <- mean(subdat$concNorm)
  subdat$sd <- sd(subdat$concNorm)
  
  #output
  dd8_norm <- rbind(dd8_norm, subdat)
}

head(dd8_norm)
```

## Data render and export
Before continuing, we need to manipulate the data. The group *Sample* becomes a factor. The sd and mean will be manually calculated per Sample. This is for exporting the data

```{r Data export}
#plot preparation
dd8_norm$Sample <- factor(dd8_norm$Sample, levels = levels(factor(dd8_norm$Sample))[1:4])
dd8_norm2 <- dd8_norm %>% group_by(Sample, .add = TRUE) %>%
              group_by(Target, .add = TRUE) %>%
              summarise(mean = mean(concNorm), SD = sd(concNorm))

#export numbers
write_xlsx(dd8_norm, path=paste(getwd(), "/ExtractedData_", ProjectName, "/", "numbersForPlot.xlsx", sep = ""))
write_xlsx(dd8_norm2, path=paste(getwd(), "/ExtractedData_", ProjectName, "/", "numbersForPlot_short.xlsx", sep = ""))

head(dd8_norm2)
```

## Statistics
We use the one-way anova, because of 4 groups. First tests will check for normal distributed data and homogeneity. Followed by data exploring statistics. 

```{r Statistics}
#check for normality
for (i in levels(dd8_norm$Sample)){
  subdat <- dd8_norm[dd8_norm$Sample == i,]
  # subdat$normality <- shapiro.test(subdat$concNorm)$p.value
  print(paste(i, shapiro.test(subdat$concNorm)$p.value, sep = ": "))
  } 
  # all tests are insignificant -> samples are normality distributed

#homogeneity of variance
leveneTest(dd8_norm$concNorm, dd8_norm$Sample, center = mean) 
  #not significant -> variances are not different=there are homogenous

#anova, one way anova
celltype<-aov(dd8_norm$concNorm ~ dd8_norm$Sample, data = dd8_norm)
summary(celltype)
#yes, there is a difference

#post-hoc, bonferroni (smaller sample size)
pairwise.t.test(dd8_norm$concNorm, dd8_norm$Sample, p.adjust.method = "bonferroni")
pairwise.t.test(dd8_norm$concNorm, dd8_norm$Sample, p.adjust.method = "BH")
pairwise.t.test(dd8_norm$concNorm, dd8_norm$Sample, p.adjust.method = "none")

#linear model (=baseline blood sample, only three tests, no correction)
summary(lm(dd8_norm$concNorm ~dd8_norm$Sample))

#t-test for comparison
test <- dd8_norm[dd8_norm$Sample %in% c("blood", "N1"),]
t.test(test$concNorm~test$Sample, var.equal = TRUE)

#statistics for the comparison with the expected 0.5 perfect heterozygosity
stats <- as.data.frame(matrix(nrow = length(levels(dd8_norm$Sample)), ncol = 4))
names(stats)<- c("Sample", "pval", "padj", "sign")

for(m in 1:length(levels(dd8_norm$Sample))){
  subdat <- dd8_norm[dd8_norm$Sample == levels(dd8_norm$Sample)[m],]
  stats$Sample[m]<- levels(dd8_norm$Sample)[m]
  stats$pval[m]<- t.test(subdat$concNorm, mu = 0.5, alternative = "less", var.equal = TRUE)$p.value
  stats$padj[m] <- p.adjust(stats$pval[m], method = "bonferroni", n = length(levels(dd8_norm$Sample)))
  
  #add pval ID
  if(stats$padj[m] < 0.001){
    stats$sign[m] <- "***"
  }else if (stats$padj[m] < 0.01){
    stats$sign[m] <- "**" 
  }else if (stats$padj[m] < 0.05){
    stats$sign[m] <- "*" 
  }else{
    stats$sign[m] <- "ns" 
  }
  # print(paste(levels(dd8_norm$Sample)[m], stats$sign[m], stats$pval[m], stats$padj[m], sep = ": "))
}
tibble(stats)


```


## Plot


```{r plot}
#plot ----
plot_norm_stat <- ggplot(dd8_norm,aes(x=Sample, y=concNorm)) +
  theme_classic(base_size=20) +
  geom_hline(yintercept = 0.5, linetype =2,size=0.5, colour = "black")+
  # geom_point(aes(col=Sample),position=position_dodge(width=0.7), size=8, alpha=0.6, shape=16)+
  geom_jitter(aes(col=Sample),size=8, alpha=0.6, shape=16, width = 0.1)+
  scale_color_manual(values=colorsForLegend[c(5,9,3,8)])+
  scale_x_discrete(labels = c("Blood", "Fibroblasts", "iPSCs", "NESCs"))+
  stat_summary(fun.data="mean_se", geom="errorbar", width=0.1, size=1)+
  geom_boxplot(aes(x=Sample, y=mean), width=0.4, size=1, colour="black")+
  labs(x="", y = "Fraction of ring chromosome", fill = element_blank()) +
  scale_y_continuous(breaks = seq(0, 0.6, by=0.1), limits = c(0,0.6))+
  annotate(geom = "text", label = "expected fraction for perfect heterozygosity", x=0.2, y = 0.5, size = 4, colour = "black", hjust=-0.03, vjust=-0.5, fontface=3)+
  theme(
    plot.subtitle = element_text(colour="black",face="italic", size = 15),
    axis.ticks = element_line(colour="black"),
    axis.text.x = element_text(colour="black", hjust=0.5, vjust=1, angle=0),
    axis.text.y = element_text(colour="black", hjust=0.5),
    legend.title = element_text(colour = "black", size = 15),
    strip.background = element_rect(size=1.5, linetype="solid", colour ="black"),
    strip.text = element_text(size = 14),
    legend.position = "none"
  )+

  annotate("text", label =paste(stats$sign)[1:2],x = stats$Sample[1:2], y = 0.58, size = 5, colour = "black", hjust=0.5, vjust=0.5)+
  annotate("text", label =paste(stats$sign)[3:4],x = stats$Sample[3:4], y = 0.58, size = 7, colour = "black", hjust=0.5, vjust=1)+ 
  annotate("text", label ="p-value:",x = 0.25, y = 0.58, size = 5, colour = "black", hjust=0, vjust=0.5, fontface=3)+
  ggtitle(paste("Mosaic state of the ring chromosome in RD_P26", sep=""),
          subtitle = paste("ddPCR (n=3), Anova: p<0.0001, bonferroni corrected post-hoc(t.test) for H0:no dif to perf.Het", sep =""))

ggsave(filename = paste("Graphs_", ProjectName, "/", PlotName, "_stats", gsub(":", "-", format(Sys.time(), "%Y%b%d_%X")), ".png", sep=""),
       height = 20, width = 25, unit="cm", limitsize = FALSE)  
plot_norm_stat 

```




